---
title: "Fungal Growth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preprocessing
```{r}
data<-read.csv('data_deal.csv')
data$log_tpm<-log(data$TPM+1)
data$log_foldchange<-log(data$FoldChange)

data_new<-data[,c(1,2,6,7)]
library(reshape2)
dmatrix<-dcast(data_new,data_new$Gene~data_new[,1],value.var = "log_foldchange")
rownames(dmatrix)<-dmatrix[,1]
dmatrix<-dmatrix[,-1]   ### become expression matrix


## Because of extra 4 experiments, it leads to some missing value in this matrix
sum(is.na(dmatrix))   ## total 7821 missing value
library(Hmisc)  
for (i in 1:36){
  dmatrix[,i]=impute(dmatrix[,i],mean)
  }                         ##impute missing value by mean

```

## Heatmap
```{r}
library(ggplot2)
library(heatmaply)
## too many genes and two many experiments, use small datasets to show the relationship
d_small=dmatrix[1:30,2:18]  
heatmaply(d_small,dendrogram = 'none')


```

##volcano map
Used to do differential expression analysis and gene selection
###Compare the differences in gene expression in different experiments.
```{r}
library(DESeq2)
count_condition<-factor(c(rep("RC1-",2),rep("RC2-",2),rep("RC3-",2),rep("RC4-",2),
                          rep("RH1-",2),rep("RH2-",2),rep("RH3-",2),rep("RH4-",2),
                          rep("YC0-",2),rep("YC1-",2),rep("TC2-",2),rep("YC3-",2),rep("YC4-",2),
                          rep("YH1-",2),rep("YH2-",2),rep("YH3-",2),rep("YH4-",2),rep("YM9-",2)))
coldata<-data.frame(row.names = colnames(dmatrix),count_condition)
##### Because DESeq require the data must be non-negative integers, so I multiply 100 and round it to get integers.
dds <- DESeqDataSetFromMatrix(countData = round(dmatrix*10), colData = coldata, design=~count_condition)
nrow(dds)
dds_out<-DESeq(dds)
res<-results(dds_out)
summary(res)
```

###Medium
```{r}
count_condition_media<-factor(c(rep('RC1',2),rep('YC1',2)))
coldata_media<-data.frame(row.names = colnames(dmatrix[,c(1,2,19,20)]),count_condition_media)
dds_media<- DESeqDataSetFromMatrix(countData = round(dmatrix[,c(1,2,19,20)]*10), colData = coldata_media, design=~count_condition_media)
dds_out_media<-DESeq(dds_media)
res_media<-results(dds_out_media)
summary(res_media)
resdata_media<- merge(as.data.frame(res_media[,]), as.data.frame(counts(dds_media, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata_media)[1] <- "Gene"
resdata_media$threshold[resdata_media$pvalue < 0.05 & resdata_media$log2FoldChange>0 ] = "up"
resdata_media$threshold[resdata_media$pvalue < 0.05 & resdata_media$log2FoldChange<0 ] = "down"
resdata_media$threshold[resdata_media$pvalue > 0.05 ] = "not_significant"

p_media<-ggplot(resdata_media,aes(x=log2FoldChange/10,y=-log10(pvalue),colour=threshold))+xlab("log2 Fold Change")+ylab("-log10 P-Value")+
  geom_point(size=4,alpha=0.6)+scale_color_manual(values =c("red","black","blue"))+ggtitle('RPMI VS. YPD')    
p_media
```

###Temperature
```{r}
count_condition_temp<-factor(c(rep('RC1',2),rep('RH1',2)))
coldata_temp<-data.frame(row.names = colnames(dmatrix[,c(1,2,9,10)]),count_condition_temp)
dds_temp<- DESeqDataSetFromMatrix(countData = round(dmatrix[,c(1,2,9,10)]*10), colData = coldata_temp, design=~count_condition_temp)
dds_out_temp<-DESeq(dds_temp)
res_temp<-results(dds_out_temp)
summary(res_temp)
resdata_temp<- merge(as.data.frame(res_temp[,]), as.data.frame(counts(dds_temp, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata_temp)[1] <- "Gene"
resdata_temp$threshold[resdata_temp$pvalue < 0.05 & resdata_temp$log2FoldChange>0 ] = "up"
resdata_temp$threshold[resdata_temp$pvalue < 0.05 & resdata_temp$log2FoldChange<0 ] = "down"
resdata_temp$threshold[resdata_temp$pvalue > 0.05 ] = "not_significant"

p_temp<-ggplot(resdata_temp,aes(x=log2FoldChange/10,y=-log10(pvalue),colour=threshold))+xlab("log2 Fold Change")+ylab("-log10 P-Value")+
  geom_point(size=4,alpha=0.6)+scale_color_manual(values =c("red","black","blue"))+ggtitle('Low Temperature VS. High Temperature')    
p_temp
```

###Time point-- one-way ANOVA
```{r}
s=c(rep('RC1',6627),rep('RC2',6627),rep('RC3',6627),rep('RC4',6627))
ss<- as.matrix(dmatrix[,c(1,2,3,4,5,6,7,8)])
dim(ss)=c(53016*1,1)
model <- aov(ss~s, data = cbind(ss,s))
summary(model)
```

##Volcano plot
```{r}
resOrdered <- res[order(res$padj), ]  
resdata <- merge(as.data.frame(resOrdered), as.data.frame(counts(dds, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata)[1] <- "Gene"
resdata$threshold[resdata$pvalue < 0.05 & resdata$log2FoldChange>0 ] = "up"
resdata$threshold[resdata$pvalue < 0.05 & resdata$log2FoldChange<0 ] = "down"
resdata$threshold[resdata$pvalue > 0.05 ] = "not_significant"

p<-ggplot(resdata,aes(x=log2FoldChange,y=-log10(pvalue),colour=threshold))+xlab("log2 Fold Change")+ylab("-log10 P-Value")+
  geom_point(size=4,alpha=0.6)+scale_color_manual(values =c("red","blue","black"))     
p
```

###Gene Selection
```{r}
resdata_filter <- subset(resdata,resdata$padj<0.05 & abs(resdata$log2FoldChange) >0)
```
###360 genes with significant expression were selected.
```{r}
summary(resdata_filter) 
write.csv(resdata_filter,'selected genes.csv')
```

## Cluster
```{r}
dcor=cor(t(dmatrix[,]),method = 'spearman')  
#more than 6000 rows, only part of it is displayed
heatmaply(dcor[1:80,1:80],ylab = 'Genes',k_col = 3,k_row = 3)

###k-means
wss<-(nrow(resdata_filter[,8:43]-1)*sum(apply(resdata_filter[,8:43],2,var)))
for (i in 2:15){
  wss[i]<- sum(kmeans(resdata_filter[,8:43],centers=i)$withinss)
}
###within-cluster sum of squares
plot(1:15, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
km<-kmeans(resdata_filter[,8:43],centers =6)
km
##Cluster result
ggplot(resdata_filter,aes(x=log2FoldChange/10,y=-log10(pvalue),color=as.character(km$cluster)))+geom_point()
```

#Q2
###PCA
```{r}
x=resdata_filter[,-c(2:7,44)]
rownames(x)<-x[,1]
x<-x[,-1]                       
pca=princomp(x,cor = T)
screeplot(pca,type = 'line',lwd=2)   ##3 principal components can be seen from the figure
score=pca$scores   ##from pca score, we can see the principle components across different experiments
write.csv(score,file = 'pca_result.csv')
pca1=pca$loadings[,1]
pca2=pca$loadings[,2]
pca3=pca$loadings[,3]
write.csv(pca$loadings[,1:3],file='pca_loading.csv')
```

#Q3
```{r}
library('Biostrings')
library(tidyverse)
gene<-read.table('H99_allorfs_promoter500nt_5mercounts.txt',header = TRUE)
rownames(gene)<-gene[,1]
gene<-gene[,-1]
```
### step1:Keep the selected genes
```{r}
left_gene=gene[rownames(x),]
```
### step2:Combine the reverse complementary segments.
```{r}
comp_dna<-data.frame(DNAStringSet(colnames(left_gene))%>%reverse%>%complement)   ##obtain reverse-complementary genes
colnames(comp_dna)='dna'

inter_dna<-intersect(DNAStringSet(colnames(left_gene)),comp_dna) 
new_couple=list()
for (i in 1:1024) {
  new_couple=c(new_couple,left_gene[,i]+left_gene[comp_dna$dna[i]])
}
new_couple_df<-data.frame(new_couple)
new_c<-t(unique(t(new_couple_df)))     ###Remove the same column
rownames(new_c)=rownames(left_gene)
```
###step3:calculate the position ratio 
###the first cluster
```{r}
x1=new_c[km$cluster==1,]

first_position=substring(colnames(x1),1,1)
second_position=substring(colnames(x1),2,2)
third_position=substring(colnames(x1),3,3)
forth_postion=substring(colnames(x1),4,4)
fifth_position=substring(colnames(x1),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x1*(first_position==x)+0)/sum(x1)
  first_per<-c(first_per,x)
}   

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x1*(second_position==x)+0)/sum(x1)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x1*(third_position==x)+0)/sum(x1)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x1*(forth_postion==x)+0)/sum(x1)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x1*(fifth_position==x)+0)/sum(x1)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x1<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x1)=base
##barplot
barplot(as.matrix(percent_position_x1),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'First Cluster',args.legend=c(x=6.8,y=1.5))
```

##the second cluster
```{r}
x2=new_c[km$cluster==2,]

first_position=substring(colnames(x2),1,1)
second_position=substring(colnames(x2),2,2)
third_position=substring(colnames(x2),3,3)
forth_postion=substring(colnames(x2),4,4)
fifth_position=substring(colnames(x2),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x2*(first_position==x)+0)/sum(x2)
  first_per<-c(first_per,x)
}   

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x2*(second_position==x)+0)/sum(x2)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x2*(third_position==x)+0)/sum(x2)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x2*(forth_postion==x)+0)/sum(x2)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x2*(fifth_position==x)+0)/sum(x2)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x2<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x2)=base
##barplot
barplot(as.matrix(percent_position_x2),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'Second Cluster',args.legend=c(x=6.8,y=1.5))
```

###the third cluster
```{r}
x3=new_c[km$cluster==3,]

first_position=substring(colnames(x3),1,1)
second_position=substring(colnames(x3),2,2)
third_position=substring(colnames(x3),3,3)
forth_postion=substring(colnames(x3),4,4)
fifth_position=substring(colnames(x3),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x3*(first_position==x)+0)/sum(x3)
  first_per<-c(first_per,x)
}   

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x3*(second_position==x)+0)/sum(x3)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x3*(third_position==x)+0)/sum(x3)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x3*(forth_postion==x)+0)/sum(x3)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x3*(fifth_position==x)+0)/sum(x3)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x3<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x3)=base
##barplot
barplot(as.matrix(percent_position_x3),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'Third Cluster',args.legend=c(x=6.8,y=1.5))
```
##the forth cluster
```{r}
x4=new_c[km$cluster==4,]
x4=x4[-42,]

first_position=substring(colnames(x4),1,1)
second_position=substring(colnames(x4),2,2)
third_position=substring(colnames(x4),3,3)
forth_postion=substring(colnames(x4),4,4)
fifth_position=substring(colnames(x4),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x4*(first_position==x)+0)/sum(x4)
  first_per<-c(first_per,x)
}   

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x4*(second_position==x)+0)/sum(x4)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x4*(third_position==x)+0)/sum(x4)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x4*(forth_postion==x)+0)/sum(x4)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x4*(fifth_position==x)+0)/sum(x4)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x4<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x4)=base
##barplot
barplot(as.matrix(percent_position_x4),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'Fourth Cluster',args.legend=c(x=6.8,y=1.5))
```

###the fifth cluster
```{r}
x5=new_c[km$cluster==5,]

first_position=substring(colnames(x5),1,1)
second_position=substring(colnames(x5),2,2)
third_position=substring(colnames(x5),3,3)
forth_postion=substring(colnames(x5),4,4)
fifth_position=substring(colnames(x5),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x5*(first_position==x)+0)/sum(x5)
  first_per<-c(first_per,x)
}   

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x5*(second_position==x)+0)/sum(x5)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x5*(third_position==x)+0)/sum(x5)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x5*(forth_postion==x)+0)/sum(x5)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x5*(fifth_position==x)+0)/sum(x5)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x5<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x5)=base
##barplot
barplot(as.matrix(percent_position_x5),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'Fifth Cluster',args.legend=c(x=6.8,y=1.5))
```

###  the sixth cluster
```{r}
x6=new_c[km$cluster==6,]

first_position=substring(colnames(x6),1,1)
second_position=substring(colnames(x6),2,2)
third_position=substring(colnames(x6),3,3)
forth_postion=substring(colnames(x6),4,4)
fifth_position=substring(colnames(x6),5,5)
## calculate percentage of first position
base<-c('A','T','G','C')

first_per=list()
for (x in base) {
  x<-sum(x6*(first_position==x)+0)/sum(x6)
  first_per<-c(first_per,x)
}  

first_per1=c(first_per[[1]],first_per[[2]],first_per[[3]],first_per[[4]])

### calculate percentage of second position
second_per=list()
for (x in base) {
  x<-sum(x6*(second_position==x)+0)/sum(x6)
  second_per<-c(second_per,x)
}
second_per1=c(second_per[[1]],second_per[[2]],second_per[[3]],second_per[[4]])

## calculate percentage of third position
third_per=list()
for (x in base) {
  x<-sum(x6*(third_position==x)+0)/sum(x6)
  third_per<-c(third_per,x)
}
third_per1=c(third_per[[1]],third_per[[2]],third_per[[3]],third_per[[4]])

## calculate percentage of forth position
forth_per=list()
for (x in base) {
  x<-sum(x6*(forth_postion==x)+0)/sum(x6)
  forth_per<-c(forth_per,x)
}

forth_per1=c(forth_per[[1]],forth_per[[2]],forth_per[[3]],forth_per[[4]])

##calculate percentage of fifth postion
fifth_per=list()
for (x in base) {
  x<-sum(x6*(fifth_position==x)+0)/sum(x6)
  fifth_per<-c(fifth_per,x)
}
fifth_per1=c(fifth_per[[1]],fifth_per[[2]],fifth_per[[3]],fifth_per[[4]])

## summary
percent_position_x6<-data.frame(first_per1,second_per1,third_per1,forth_per1,fifth_per1)
rownames(percent_position_x6)=base
##barplot
barplot(as.matrix(percent_position_x6),names.arg = c('1','2','3','4','5'),
        col = rep(c('lightgreen','lightyellow','lightblue','blue'),5),legend.text = TRUE,main = 'Sixth Cluster',args.legend=c(x=6.8,y=1.5))
```

### prediction analysis
```{r}
x1_mean=apply(x1,2,mean)
x2_mean=apply(x2,2,mean)
x3_mean=apply(x3,2,mean)
x4_mean=apply(x4,2,mean)
x5_mean=apply(x5,2,mean)
x6_mean=apply(x6,2,mean)
```